Envelope Estimation using Geometric Properties of a Discrete Real Signal

### 2020-12-05 Papers/01_Pulses_Envelope/PulsesEnvelope.tex
grammar checked

% \usepackage{bm} % bold & upright inside math env
% \usepackage{hyperref}
% \usepackage{hypcap}
\DeclareMathOperator{\sign}{sign}
\bibliography{bibli}
% \hypersetup{
%   colorlinks=true, 
%   allcolors=black,
%   pdfauthor={Carlos Henrique Tarjano Santos},
%   pdftitle={Robust Digital Envelope Estimation Via Geometric Properties of an Arbitrary Real Signal}
% }
\title{Envelope Estimation using Geometric Properties of a Discrete Real Signal}
\author[1]{Carlos Henrique Tarjano Santos}
\author[2]{Valdecy Pereira}
\affil[1]{(corresponding author, carlostarjano@id.uff.br) Department of Production Engineering, Universidade Federal Fluminense, Rua Passo da Pátria, 156, Campus Praia Vermelha, Bloco D - sala 309, São Domingos, Niterói, RJ, Brasil, CEP: 24.210-240}
\affil[2]{Department of Production Engineering, Universidade Federal Fluminense, Rua Passo da Pátria, 156, Campus Praia Vermelha, Bloco D - sala 309, São Domingos, Niterói, RJ, Brasil, CEP: 24.210-240}

@@

\begin{abstract}
% Background
Despite being an elusive concept, the temporal amplitude envelope of a signal is essential for its complete characterization, being the primary information-carrying medium in spoken voice and telecommunications, for example. Intuitively, the temporal envelope can be understood as a slow varying function that multiplies the signal, being responsible for its outer shape. 
Envelope detection techniques have applications in areas like health, sound classification and synthesis, seismology and speech recognition. 
% Problem | Objectives
Nevertheless, a general approach to digital envelope detection of signals with rich spectral content doesn't exist, as most methods involve manual intervention, in the form of filter design, smoothing, as well as other specific design choices, based on a priori knowledge of the nature of the specific waves under investigation.
To address this problem, we propose a framework that uses intrinsic characteristics of a signal to estimate its envelope, eliminating the necessity of parameter tuning.
% Methods | Procedure | Approach
The approach here described draws inspiration from geometric concepts to estimate the temporal envelope of an arbitrary signal; specifically, a new measure of discrete curvature is used to obtain the average radius of curvature of a discrete wave, that will as a threshold to identify the wave's samples that are part of the envelope, in a manner similar to the alpha-shapes approach to the identification of the concave hull of a set of points.
% Results | Findings | Product
We provide visualizations of the envelope extracted via the algorithm for various real world signals, with very diverse characteristics, such as voice, spoken and sang, and pitched and non pitched musical instruments, and discuss some approaches to assess the quality of the obtained envelopes. The algorithm compares favourably with classic envelope detection techniques based on filtering and the Hilbert Transform. A Python module, installable via the Python Package Index (PyPI), with a pure Python and a specialized C++ implementation for 64 bit Windows systems, and interactive visualizations of envelopes for a diverse range of digital waves, as well as the source code for the Python implementation, are available online.
% Conclusions | Implications
Besides the most direct applications of this work to audio classification and synthesis, we foresee impact in compression techniques and machine learning approaches to audio. The discrete curvature definition presented could also be extended to three-dimensional settings, to improve shape detection algorithms based on alpha-shapes.
\end{abstract}

{\bf Keywords:} DSP, alpha-shapes, envelope detection, discrete curvature estimation, demodulation

\section{Introduction}
Envelope detection, also known as demodulation, is ubiquitous in both analogue and digital signal processing \parencite{2011CaetanoImproved}. Nevertheless, the literature in this area is very fragmented \parencite{2017LyonsDigital}. Besides, most envelope detection techniques are designed to account for very specific settings, like pure sinusoids with moderate noise content, a limitation that excludes most physical signals, as is the case of recorded sound, for example; that limitation arises in part due to the lack of a strict mathematical definition of a temporal envelope \parencite{2013Mengempirical}.

@@

One direct way of defining the shape of a set of points in $ \mathbb{R}^2 $ is using the convex hull: as defined in \textcite{2008BergComputational}, the convex hull of a set of points $ S $ in two dimensional space is the smallest convex subset of the plane that contains all the points in $ S $. Convex hulls are unique, in the sense that a set of points define one and only one convex hull.

Generalizing that definition, \textcite{1983Edelsbrunnershape} introduced the concept of alpha-shapes, that can be seen as a Concave Hull; a mathematically well-defined generalization of the Convex Hull of a finite set of points, closely related to the Delaunay triangulation and Voronoi diagrams of those points.

The intuition behind the alpha shapes algorithm as that only points that can be touched by a circle of radius $ \alpha $ coming from an infinite distance towards the set of points of interest are part of the frontier, that is, of the Concave Hull of this set.

Once an $ \alpha $ parameter is defined alpha shapes conserve the uniqueness property, besides reducing to the Convex Hull when $ \alpha = \infty $. Consider, for example, the artificially generate discrete wave in figure \ref{fig:Hulls}
below, where the links between consecutive samples were deemphasised to highlight the "set of points" nature of a signal: The Convex Hull and a Concave Hull, in the form of an alpha shape with a particular value of $ \alpha $ are illustrated.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/00ConvexHull.pdf}
  \caption{A discrete signal interpreted as a set of points in the Cartesian plane, and the Convex Hull and Concave Hull at $ \alpha=300 $ defined by it}
  \label{fig:Hulls}
\end{figure}

@@

Following this path, we present a general approach to envelope detection, exploiting the intrinsic characteristics of a generic, spectrally complex wave, in order to avoid as much as possible the need for manual intervention or parameter tuning, such as the choice of $ \alpha $.

% TODO <|<|<|<|<|

While the robustness of the proposed approach allows it to be used as a plug-in replacement for many methods encountered in the literature, we feel that it would be particularly useful for sound synthesis.

The envelope is shown to add complexity to the spectral representation of a wave \parencite{2019TarjanoNeuro}, and an accurate description of the envelope, while describing the evolution of the instantaneous amplitude of a signal in time, would also greatly simplify further spectral analysis. Moreover, the algorithm developed in this work naturally divides a signal into its pseudo-cycles, pinpointing them in the time domain, providing the building blocks for the reconstruction of the fine structure of the wave.

The rest of this paper is divided as follows: After explaining and defining the various entities to be used subsequently, of which the concept of frontiers is arguably the most important, in the methodology section, we dedicate a section to present a new discrete curvature estimation approach, and how it was used to identify the frontiers of a discrete wave.

We then proceed to illustrate the results of the application of the algorithm in a set of six diverse sound waves, ranging from voice to unpitched instruments comparing, afterwards, its performance with that of classic digital envelope estimation algorithms, both from a visual and numerical standpoint.

We then discuss the implications of the work, emphasizing the focus on the construction of new interpretations of the relationship between time domain and frequency domain sound descriptions, and our vision of the impact of this methodology in the field of sound synthesis, providing directions for future developments.

\section{Methodology}
% Overview of methods

@@

The interest in the theory of envelope detection arose in the analogue domain, with the widespread adoption of radio communications, where the problem is typically restricted to simple sinusoids, often with well defined frequencies \parencite{2011TurnerDemodulation}.

In this context, the most mathematically sound definition of an envelope involves the representation of its underlying wave as an analytic signal, introduced by Gabor in 1946 \parencite{2007HahnHistory}. In his work, \textcite{1946GaborTheory} applies the then relatively new mathematical machinery of the quantum mechanics to unify the time and frequency domain representations of a wave, showing how the Hilbert transform could be applied to a real signal in order to obtain an equivalent complex signal, later known as the analytic signal.

@@

@@

In general, the problem of envelope detection, or demodulation, can be interpreted as the task of, given a continuous wave $ W(t) $, decomposing it in two components $ E(t) $ and $ C(t) $, such that $ W(t) = E(t) C(t) $ \parencite{2011TurnerDemodulation}. 
$ E(t) $ represents the slow varying part of the wave, also known as the (temporal) envelope, modulator component, or amplitude modulation (AM) of $ W $, while $ C(t) $ models its fast varying part, called throughout the literature as its (temporal) fine structure, carrier component, or frequency modulation (FM).

In the case of broadband signals, the problem of envelope detection of an arbitrary wave is ill posed, in the sense that an infinite number of pairs of $ E(t) $ and $ C(t) $ can result in the same specific $ W(t) $ \parencite{2011TurnerDemodulation, 2013Mengempirical, 1996Loughlinamplitude}.

@@

\begin{equation} \label{eq:Envelope}
  \begin{align}
    & \textbf{w}, \textbf{e}, \textbf{c} \in \mathbb{R}^n \\
    & \textbf{w} = (w_0, w_1, \cdots, w_{n-1}) \\
    & \textbf{e} = (e_0, e_1, \cdots, e_{n-1}) \\
    & \textbf{c} = (c_0, c_1, \cdots, c_{n-1}) \\
    & \textbf{w} = \textbf{e} \odot \textbf{c} \\
    & w_i = e_i c_i \quad \forall \quad i, 0 \le i \le n-1
  \end{align}
\end{equation}

@@

% TODO <|<|<|<|<| we will assume that this assumption holds for quasi periodic signals: that is, the absolute maximum of each pseudo cycle is equal to the value of the envelope at that point, the other local extrema in that same cycle are the ones that should be filtered out.

% TODO <|<|<|<|<| We lack a indicator to evaluate the quality of the envelope, once one is identified, but figure suggests that, once the envelope is removed from w, the wave should then be confined between the interval $ \{-1, 1\}$.

Image \ref{fig:Envelope}, along with the Concave Hull illustrated in figure \ref{fig:Hulls}, helps forming the intuition behind the method developed in this work: the idea is to identify the local extrema that touch the envelope (or, to put in another way, the local extrema in relation to the whole wave that are global extrema inside their own pseudo cycles), filtering out those who don't; in this way, we're defining the envelope itself at those points.

@@

In the next section such a geometric representation will be formulated, along with the necessary tool to estimate the average curvature of the wave and, consequently, the appropriate radius of the circle.

% TODO <|<|<|<|<|

@@

P is not fit for geometric interpretation yet, since the abscissa and ordinate of his orthogonal coordinate system have different units: in the vertical axis we have a unit related to the instantaneous intensity of the wave, while in the horizontal axis we have the index at which each extremum occurs, ultimately a time related unit.

@@

We are interested in finding a coordinate system where a pulse's amplitude $ \lvert w_j \lvert $ and length would define, in average, a square.

@@

The drawback is that we would lose the direct link between P and \textbf{w} that arises from the fact that the abscissas of the points in P are indices $ i $ of \textbf{w}: we choose instead to leave the abscissa intact and multiply the now unitless vertical basis axis by the average of the difference between the abscissa of a point and the abscissa of the immediately posterior. In this way, both axes will have unit $ i $, and the relation is preserved, making it easier to recover the envelope points later.

@@

\begin{equation} \label{eq:Basis}
\begin{align}
\mathbf{b}^\prime_x &= \mathbf{b}_x  \\
\mathbf{b}^\prime_y &= \frac{\left( \frac{i_{m-1} - i_0}{m-1} \right)}{\left( \frac{\sum_{j=0}^{m-1} \lvert w_j \lvert}{m} \right)} \mathbf{b}_y 
\end{align}
\end{equation}

@@

\begin{equation} \label{eq:Vectors}
\begin{align}
\text{V} = \{ v_0, v_1, v_{m-2} \} \quad \text{where} \quad v_k = P_{j+1} - P_j \quad \forall \quad 0 \le k \le m-2
\end{align}
\end{equation}

@@

Translating the explanation of alpha-shapes in \textcite{1994EdelsbrunnerThree} to our context, we can say that the points that are part of the envelope are those touched by a circle of a given radius $ r $, coming from $ y \to \infty $ towards the signal, that is not allowed to contain any point of the signal. % TODO refine

Intuitively, one can picture a circle being rolled above the signal, and marking the points it touches as envelope points.

@@

% Concerning the last method, two approaches to fit a polynomial to a generic wave are readily available: the least mean squares approach, that seeks to minimize the dependent variable errors and the total least squares problem, that treats both variables symmetrically \parencite{1980GolubAnalysis}.

% The geometric, symmetrical nature of the problem excludes the more computationally economic least mean squares approach, however, leaving us with the total least squares, for which no general closed-form solutions are available \parencite{2007MarkovskyOverview}; this leads us to turn our attention to direct approaches of curvature estimation.

Concerning direct methods \textcite{2014CarrollSurvey}, for example, derive three such definitions based on the approximation of a circle by an inscribed, centred and circumscribed polygon. In the context of tridimentional meshes, \textcite{2016VasaMultivariate} evaluate a range of existing estimators from a multivariate point of view.

@@

\begin{equation} \label{eq:Radius}
% \begin{gather}
r = \frac{\sum\limits_{k=0}^{m-2} \left( \frac{v_{k,x} \sqrt{v_{k,x}^2 + v_{k,y}^2}}{ v_{k,y}} \right)}{m-2}
% \end{gather}
\end{equation}

We need now to construct an algorithm, using the radius just obtained, to identify the points that belong to the envelope.
Generally, the algorithm for the alpha-shapes approach resorts to the construction of the Delaunay triangulation of the set of points, that is latter filtered. We can adopt a more straightforward method, as our points have a defined structure:

@@

\begin{equation} \label{eq:InEnvelope}
\begin{align}
& P_a \in \text{E} \leftrightarrow (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 > r^2, \\
& \forall \ P_b, P_c \mid (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 = (P_{c,x} - x_o)^2 + (P_{c,y} - y_o)^2 = r^2, \\
& P_a, P_b, P_c \in \text{P}, \quad P_a \ne P_b \ne P_c \quad \text{and} \quad x_o, y_o \in \mathbb{R}, \quad 0 \le y_o < +\infty
\end{align}
\end{equation}

@@

The set E obtained via algorithm \ref{FindEnvelope} is a set of points. The most straightforward way to transform those points into the vector $ \textbf{e} \in \mathbb{R}^n $ is via linear interpolation, the approach used in this work. 
is worthy noting, however, that many other procedures are available, both for the interpolation and the smooth approximation of points. We deliberately choose linear interpolation to avoid blurrying the results of the method.

\section{Results}
% Revisit purpose and methods
% Overview of results

@@

There are no agreed upon objective metrics to assess the quality of an envelope. Recalling figure \ref{fig:Envelope}, however, and the accompanying discussion, an intuitive way is to consider the carrier wave \textbf{c}, and how well it seems to be bounded in amplitude in the interval $ \{-1, 1\} $. To that end, with provide the follow figures of original discrete waves \textbf{w}, their envelope \textbf{e}, and the inferred carrier, obtained after dividing, element wise, the original wave by its envelope.

@@

Due to an assortment of factors, such as the unconventional approach presented in this work and the intrinsic dificult in visualizing discrete waves given their high dimensionality, the authors deemed helpful to provide a companion website with interactive visualizations.

An implementation of the method in pure Python, fully functional, can also be found at the repository dedicated to this work as well as a Python module available at the Python Package Index (PyPI) and conveniently instalabbe via the command `pip install signal-envelope`. Besides, those working under Windows 64 bit can benefit from a specialized dll coded in C++ that is automatically used within this module.

@@

The first one states that if a signal is bounded in magnitude, then its envelope should be, as well. Our algorithm complies with this requisite, as the envelope is composed of selected samples from the original signal itself;

@@

Nevertheless, insight can be gained from a comparison of the results of the method here proposed with some of the most common envelope extraction algorithms, as can be seen in figure \ref{fig:Comparison}.

\begin{figure}[ht!]
  \centering
    \includegraphics[width=0.8\linewidth]{12Comparison.pdf}
  \caption{Comparison of the algorithm present in this work with the most common methods of digital envelope identification.}
  \label{fig:Comparison}
\end{figure}

For a recording of a guitar sound, in which a bend is performed, we contrast the proposed algorithm with four common approaches of varying complexity: simple smoothing of the rectified original wave, low pass filtering followed by rectification, peak identification and subsequent smoothing, and the extraction of the envelope via Hilbert transform, as presented earlier in the paper, with a previous smoothing of the original wave to avoid the artifacts shown in figure \ref{fig:AnalyticSignal}.

All those methods demanded careful choice of parameters: in the case of filtering, a cut-off frequency of approximately 88 Hz was used, consisting of one additional parameter to be tuned. For all the other traditional approaches, the Savitzky-Golay smoothing algorithm was used, and the window size was the parameter to be defined.

In the case of the proposed algorithm, the smooth temporal envelope was obtained merging both positive and negative frontiers, the result being smoothed with the same procedure described. Figure \ref{fig:FrontiersToEnvelope} illustrates the results.

\begin{figure}[ht!]
  \centering
    \includegraphics[width=0.8\linewidth]{13FrontiersToEnvelope.pdf}
  \caption{Original positive and negative frontiers in dark grey and their mirrored versions. The temporal envelope, in black, is obtained by merging and approximating both frontiers.}
  \label{fig:FrontiersToEnvelope}
\end{figure}

Table \ref{table:Comparison} presents a computational comparison between the techniques. The times are for custom implementations of those methods in the Python programming language. The least mean squares result was calculated by halving the envelopes presented in figure \ref{fig:Comparison} and comparing them with the rectified absolute wave, so as to avoid the potential influence of an estimated frequency.

We can see that, while the algorithm here presented is the most accurate, it's also the most computationally intensive. The connection between alpha-shapes and Delaunay triangulation can be explored in future works as a means to improve the efficiency of the algorithm.

>>>>

% \usepackage{hyperref}
% \usepackage{hypcap}
% \hypersetup{
%   colorlinks=true, 
%   allcolors=black,
%   pdfauthor={Carlos Henrique Tarjano Santos},
%   pdftitle={Robust Digital Envelope Estimation Via Geometric Properties of an Arbitrary Real Signal}
% }

\DeclareMathOperator{\sign}{sign}
\bibliography{bibli}

\title{Envelope Estimation using Geometric Properties of a Discrete Real Signal}
\author[1]{Carlos Henrique Tarjano Santos}
\author[2]{Valdecy Pereira}
\affil[1]{(corresponding author, carlostarjano@id.uff.br) Department of Production Engineering, Universidade Federal Fluminense, Rua Passo da Pátria, 156, Campus Praia Vermelha, Bloco D - sala 309, São Domingos, Niterói, RJ, Brasil, CEP: 24.210-240}
\affil[2]{Department of Production Engineering, Universidade Federal Fluminense, Rua Passo da Pátria, 156, Campus Praia Vermelha, Bloco D - sala 309, São Domingos, Niterói, RJ, Brasil, CEP: 24.210-240}

@@

\begin{abstract}
% Background
Despite being an elusive concept, the temporal amplitude envelope of a signal is essential for its complete characterization, being the primary information-carrying medium in spoken voice and telecommunications, for example. Intuitively, the temporal envelope can be understood as a slowly varying function that multiplies the signal, being responsible for its outer shape. 
Envelope detection techniques have applications in areas like health, sound classification and synthesis, seismology and speech recognition. 
% Problem | Objectives
Nevertheless, a general approach to digital envelope detection of signals with rich spectral content doesn't exist, as most methods involve manual intervention, in the form of filter design, smoothing, as well as other specific design choices, based on prior knowledge of the nature of the specific waves under investigation.
To address this problem, we propose a framework that uses intrinsic characteristics of a signal to estimate its envelope, eliminating the necessity of parameter tuning.
% Methods | Procedure | Approach
The approach here described draws inspiration from geometric concepts to estimate the temporal envelope of an arbitrary signal; specifically, a new measure of discrete curvature is used to obtain the average radius of curvature of a discrete wave, that will as a threshold to identify the wave's samples that are part of the envelope, in a manner similar to the alpha shapes approach to the identification of the concave hull of a set of points.
% Results | Findings | Product
The algorithm compares favourably with classic envelope detection techniques based on smoothing, filtering and the Hilbert Transform, besides being physically plausible.
We provide visualizations of the envelope extracted via the algorithm for various real-world signals, with very diverse characteristics, such as voice, spoken and sang, and pitched and non-pitched musical instruments, and discuss some approaches to assess the quality of the obtained envelopes.
A Python module, installable via the Python Package Index (PyPI), with a pure Python and a specialized C++ implementation for 64-bit Windows systems, interactive visualizations of envelopes for a diverse range of digital waves, as well as the source code for the Python implementation, are available online at \textcite{2020TarjanoEnvelope}.
% Conclusions | Implications
Besides the most direct applications of this work to audio classification and synthesis, we foresee impact in compression techniques and machine learning approaches to audio. The discrete curvature definition presented could also be extended to three-dimensional settings, to improve shape detection algorithms based on alpha shapes.
\end{abstract}

{\bf Keywords:} DSP, alpha shapes, envelope detection, discrete curvature estimation, demodulation

\section{Introduction}
Envelope detection, also known as demodulation, is ubiquitous in both analogue and digital signal processing \parencite{2011CaetanoImproved}. Nevertheless, the literature in this area is fragmented \parencite{2017LyonsDigital}. Besides, most envelope detection techniques are designed to account for very specific settings, like pure sinusoids with moderate noise content, a limitation that excludes most physical signals, as is the case of recorded sound, for example; that limitation arises in part due to the lack of a strict mathematical definition of a temporal envelope \parencite{2013Mengempirical}.

@@

One direct way of defining the shape of a set of points in $ \mathbb{R}^2 $ is using the convex hull: as defined in \textcite{2008BergComputational}, the convex hull of a set of points $ S $ in two-dimensional space is the smallest convex subset of the plane that contains all the points in $ S $. Convex hulls are unique, in the sense that a set of points define one and only one convex hull.

Generalizing that definition, \textcite{1983Edelsbrunnershape} introduced the concept of alpha shapes, that can be seen as a Concave Hull; a mathematically well-defined generalization of the Convex Hull of a finite set of points, closely related to the Delaunay triangulation and Voronoi diagrams of those points.

The intuition behind the alpha shapes algorithm is that only points that can be touched by a circle of radius $ \alpha $ coming from an infinite distance towards the set of points of interest are part of the frontier, that is, of the Concave Hull of this set.

Once an $ \alpha $ parameter is defined alpha shapes conserve the uniqueness property, besides reducing to the Convex Hull when $ \alpha = \infty $. Consider, for example, the artificially generated discrete wave in figure \ref{fig:Hulls}
below, where the links between consecutive samples were deemphasized to highlight the “set of points” nature of a signal: The Convex Hull and a Concave Hull, in the form of an alpha shape with a particular value of $ \alpha $ are illustrated.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/00ConvexHull.pdf}
  \caption{A discrete signal interpreted as a set of points in the Cartesian plane, and the Convex Hull and Concave Hull at $ \alpha=300 $ defined by it.}
  \label{fig:Hulls}
\end{figure}

@@

Following this path, we present a general approach to envelope detection, exploiting the intrinsic characteristics of a generic, spectrally complex wave, in order to completely avoid the need for manual intervention or parameter tuning.

The rest of this paper is divided as follows: 
We start by trying to define as specifically as possible what a temporal envelope is, especially in the context of this work, providing a mathematical definition, in terms of the Hilbert transform, for the case of a narrowband signal.

We proceed to present the representation of a discrete wave used in this work, largely based on the concept of dividing the wave into its constituent pulses, and how it can simplify the algorithm, reducing the high dimensionality generally present by digital representations of waves.

Next, we review methods for the estimation of the curvature of a discrete wave, after which we present our own; this method will then be used to estimate the equivalent circle, that will be used to define the envelope of the wave.

In the results section, we present some illustrations of the envelopes obtained with our method, comparing both accuracy and running time with the most common algorithms encountered in the literature. We also suggest an alternative use of the algorithm to identify the negative and positive frontiers of a wave, a kind of partial envelope considering only the positive and negative pulses of a wave, respectively.

We follow with some suggestions about how the approach here presented can be useful beyond envelope detection, exemplifying with some applications, such as the extractions of the average waveform of a digital wave, concluding with some commentary on potential future developments of the research here presented.

\section{Methodology}
After defining what will be considered an envelope in the context of this work, in this section we explore some simplifications that can be applied to a discrete wave, defining the representation that will be used in the subsequent part of the work.

We next explore the concept of discrete curvature, putting forth our own method, tailored to the definition of the equivalent circle of a wave; we end this section by showing how this circle can be used to identify the envelope of a discrete wave, in a similar approach to that of the alpha shapes theory, but taking advantage of some unique structure present in discrete waves.

@@

The interest in the theory of envelope detection arose in the analogue domain, with the widespread adoption of radio communications, where the problem is typically restricted to simple sinusoids, often with well-defined frequencies \parencite{2011TurnerDemodulation}.

In this context, the most mathematically sound definition of an envelope involves the representation of its underlying wave as an analytic signal, introduced by Gabor in 1946 \parencite{2007HahnHistory}. In his work, \textcite{1946GaborTheory} applies the then relatively new mathematical machinery of the quantum mechanics to unify the time and frequency-domain representations of a wave, showing how the Hilbert transform could be applied to a real signal in order to obtain an equivalent complex signal, later known as the analytic signal.

@@

%%%%%%%%%%%%%%%%%

@@

In general, the problem of envelope detection, or demodulation, can be interpreted as the task of, given a continuous wave $ W(t) $, decomposing it in two components $ E(t) $ and $ C(t) $, such that $ W(t) = E(t) C(t) $ \parencite{2011TurnerDemodulation}.

$ E(t) $ represents the slow varying part of the wave, also known as the (temporal) envelope, modulator component, or amplitude modulation (AM) of $ W $, while $ C(t) $ models its fast varying part, called throughout the literature as its (temporal) fine structure, carrier component, or frequency modulation (FM).

In the case of broadband signals, the problem of envelope detection of an arbitrary wave is ill-posed, in the sense that an infinite number of pairs of $ E(t) $ and $ C(t) $ can result in the same specific $ W(t) $ \parencite{2011TurnerDemodulation, 2013Mengempirical, 1996Loughlinamplitude}.

@@

\begin{align} \label{eq:Envelope}
  & \textbf{w}, \textbf{e}, \textbf{c} \in \mathbb{R}^n \\
  & \textbf{w} = (w_0, w_1, \cdots, w_{n-1}) \\
  & \textbf{e} = (e_0, e_1, \cdots, e_{n-1}) \\
  & \textbf{c} = (c_0, c_1, \cdots, c_{n-1}) \\
  & \textbf{w} = \textbf{e} \odot \textbf{c} \\
  & w_i = e_i c_i \quad \forall \quad i, 0 \le i \le n-1
\end{align}

@@

Image \ref{fig:Envelope}, along with the Concave Hull illustrated in figure \ref{fig:Hulls}, helps forming the intuition behind the method developed in this work: the idea is to identify the local extrema that touch the envelope (or, to put in another way, the local extrema in relation to the whole wave that are global extrema inside their own pseudo-cycles), filtering out those who don't; in this way, we're defining the envelope itself at those points.

@@

In the next section, such a geometric representation will be formulated, along with the necessary tool to estimate the average curvature of the wave and, consequently, the appropriate radius of the circle.

@@

P is not fit for geometric interpretation yet, since the abscissa and ordinate of his orthogonal coordinate system have different units: in the vertical axis, we have a unit related to the instantaneous intensity of the wave, while in the horizontal axis we have the index at which each extremum occurs, ultimately a time-related unit.

@@

We are interested in finding a coordinate system where a pulse's amplitude $ \lvert w_j \lvert $ and length would define, on average, a square.

@@

The drawback is that we would lose the direct link between P and \textbf{w} that arises from the fact that the abscissas of the points in P are indices $ i $ of \textbf{w}: we choose instead to leave the abscissa intact and multiply the now unitless vertical basis axis by the average of the difference between the abscissa of a point and the abscissa of the immediately posterior. In this way, both axes will have unit $ i $, and the relationship is preserved, making it easier to recover the envelope points later.

@@

\begin{align} \label{eq:Basis}
\mathbf{b}^\prime_x &= \mathbf{b}_x  \\
\mathbf{b}^\prime_y &= \frac{\left( \frac{i_{m-1} - i_0}{m-1} \right)}{\left( \frac{\sum_{j=0}^{m-1} \lvert w_j \lvert}{m} \right)} \mathbf{b}_y 
\end{align}

@@

\begin{equation} \label{eq:Vectors}
\text{V} = \{ v_0, v_1, v_{m-2} \} \quad \text{where} \quad v_k = P_{j+1} - P_j \quad \forall \quad 0 \le k \le m-2
\end{equation}

@@

Translating the explanation of alpha shapes in \textcite{1994EdelsbrunnerThree} to our context, we can say that the points that are part of the envelope are those touched by a circle of a given radius $ r $, coming from $ y \to \infty $ towards the signal, that is not allowed to contain any point of the signal; Intuitively, one can picture a circle being rolled above the signal, and marking the points it touches as envelope points.

@@

Concerning direct methods \textcite{2014CarrollSurvey}, for example, derive three such definitions based on the approximation of a circle by an inscribed, centered and circumscribed polygon. In the context of tridimensional meshes, \textcite{2016VasaMultivariate} evaluate a range of existing estimators from a multivariate point of view.

@@

\begin{equation} \label{eq:Radius}
r = \frac{\sum\limits_{k=0}^{m-2} \left( \frac{v_{k,x} \sqrt{v_{k,x}^2 + v_{k,y}^2}}{ v_{k,y}} \right)}{m-2}
\end{equation}

We need now to construct an algorithm, using the radius just obtained, to identify the points that belong to the envelope.
Generally, the algorithm for the alpha shapes approach resorts to the construction of the Delaunay triangulation of the set of points, that is latter filtered. We can adopt a more straightforward method, as our points have a defined structure:

@@

\begin{align} \label{eq:InEnvelope}
& P_a \in \text{E} \leftrightarrow (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 > r^2, \\
& \forall \ P_b, P_c \mid (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 = (P_{c,x} - x_o)^2 + (P_{c,y} - y_o)^2 = r^2, \\
& P_a, P_b, P_c \in \text{P}, \quad P_a \ne P_b \ne P_c \quad \text{and} \quad x_o, y_o \in \mathbb{R}, \quad 0 \le y_o < +\infty
\end{align}

@@

The set E obtained via algorithm \ref{FindEnvelope} is a set of points. The most straightforward way to transform those points into the vector $ \textbf{e} \in \mathbb{R}^n $ is via linear interpolation, the approach used in this work. 
It is worth noting, however, that many other procedures are available, both for the interpolation and the smooth approximation of points. We deliberately choose linear interpolation to avoid blurring the results of the method.

\section{Results}
Given the lack of a precise definition of what an envelope should be, we begin this section suggesting an empirical metric, based on the behaviour of the carrier wave \textbf{c}. We also comment on some metrics presented in the literature, and compare our algorithm with traditional demodulation approaches, both in relation to execution time and accuracy; to this last end, a numerical indicator is introduced.

We also illustrate how the method here presented can be used to isolate not only a single envelope of the wave, but the superior and inferior envelopes as well, that we call frontiers, ending the section with a suggestion on how the proposed algorithm can be useful beyond envelope detection, by identifying the approximate locations of the pseudo-cycles in a quasi-periodic wave.

Due to an assortment of factors, such as the unconventional approach presented in this work and the intrinsic difficulty in visualizing discrete waves given their high dimensionality, the authors deemed helpful to provide a companion website with interactive visualizations \parencite{2020TarjanoEnvelope}.

An implementation of the method in pure Python, fully functional, can also be found at the repository dedicated to this work as well as a Python module available at the Python Package Index (PyPI) and conveniently installable via the command `pip install signal-envelope`. Besides, those working under Windows 64-bit can benefit from a specialized dynamic library coded in C++ that is automatically used within this module.

@@

There are no agreed-upon objective metrics to assess the quality of an envelope. Recalling figure \ref{fig:Envelope}, however, and the accompanying discussion, an intuitive way is to consider the carrier wave \textbf{c}, and how well it seems to be bounded in amplitude in the interval $ \{-1, 1\} $. To that end, we provide the following figures of original discrete waves \textbf{w}, their envelope \textbf{e}, and the inferred carrier, obtained after dividing, element-wise, the original wave by its envelope.

From figure \ref{fig:spoken_voice} it can be seen that the algorithm performs well under the carrier metric. This is especially remarkable for the case of spoken voice, a notorious complex sound with fast changes in pitch and high noise content.

These characteristics are in stark contrast with the sang voice in figure \ref{fig:singer_voice}, where a recording of an alto singer is shown; there, the wave is almost periodic, with a stable waveform.

@@

\begin{figure}[ht!]
  \centering
    \includegraphics{images/06Graph_spoken_voice.pdf}
  \caption{Signal, envelope and carrier for a record of male voice uttering the word "amazing".}
  \label{fig:spoken_voice}
\end{figure}

\begin{figure}[ht!]
  \centering
    \includegraphics{images/06Graph_alto.pdf}
  \caption{Signal, envelope and carrier for a record of an alto singer.}
  \label{fig:singer_voice}
\end{figure}

@@

The first one states that if a signal is bounded in magnitude, then its envelope should be, as well. Our algorithm complies with this requisite, as the envelope is composed of selected samples from the original signal itself.

@@

Nevertheless, insight can be gained from a comparison of the results of the method here proposed with some of the most common envelope extraction algorithms. Specifically, we compare the present method with the following approaches:

\begin{enumerate}
\item Smoothing: This is a relatively simple procedure that, applied to the absolute values of a signal, can provide an approximation of the variation of its amplitude in time, by smoothing the movement of the carrier. We use the Savitzky-Golay algorithm \parencite{1964SavitzkySmoothing} with a window width of 3001 samples and a cubic polynomial.
\item Filtering: The absolute value of the wave is filtered using a digital implementation of the Butterworth \parencite{butterworth1930theory} low pass filter. We use an order 2 filter with a cut off frequency of 10 Hz.
\item Hilbert Transform: We apply the Hilbert transform to the absolute values of the wave after filtering the original signal with a Butterworth filter with a cut off frequency of 100 Hz, using the absolute value of the analytic signal to obtain the envelope.
\end{enumerate}

In \ref{fig:Comparison_sinusoid} we illustrate the envelope extracted by the conventional algorithms, contrasting it with the result of the method presented in this work, for a simple sinusoid. Note that none of the conventional algorithms complies with the third condition proposed in \textcite{1996Loughlinamplitude} that the envelope of a periodic wave should be a straight line.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/07Comparison_sinusoid.pdf}
  \caption{Comparison of envelope detection algorithms for a simple sinusoid}
  \label{fig:Comparison_sinusoid}
\end{figure}

For a recording of the key 33 of a grand piano, where parts of the sound are highly percussive and noisy, the Hilbert approach presents considerable oscillations, as can be seen in figure \ref{fig:Comparison_piano}.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/07Comparison_piano.pdf}
  \caption{Comparison of envelope detection algorithms for the recording of the key 33 of a grand piano}
  \label{fig:Comparison_piano}
\end{figure}

In figure \ref{fig:Comparison_soprano} all envelopes exhibit a similar movement along the samples, but the ones generated by traditional algorithms are undershooting the original wave.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/07Comparison_soprano.pdf}
  \caption{Comparison of envelope detection algorithms for the recording of the vocalization of a soprano singer}
  \label{fig:Comparison_soprano}
\end{figure}

All the traditional methods demanded careful choice of parameters. In the case of the Hilbert transform, pre-processing in the form of filtering was applied. That is in contrast with our method, that organically tunes itself, in the form of the automatic choice of the radius of the circle representing the average discrete curvature of the discrete signal.

Table \ref{table:Error_comparison} presents a comparison of the average errors of the techniques. The average error was calculated by the average of the squared distance of the points in the envelope divided by two and the absolute value of the wave. The rationale is that the values of the envelope divide by two should pass near the average of the absolute values of the wave.

We can see from the table \ref{table:Error_comparison} that the method presented here consistently outperforms the traditional approaches for all waves tested, yielding an error 33\% smaller, on average, than the other methods. It is also worth noting that the errors of the traditional methods are very similar between them.
